---
title: "Prediction model for permanent hypoparathyroidism"
author: "Carolien C.H.M. Maas, Erasmus MC, Rotterdam, The Netherlands"
date: "July 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages, functions, and data
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
seed <- 1
set.seed(seed)

# load libraries
library(haven)
library(dplyr)
library(mice)
library(rms)
```

# Load data
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
file.path <- "Z:/Project Predict Hypoparathyroidism/Development/"
original.data <- read.csv(paste0(file.path, "Data/Data.csv"), 
                 stringsAsFactors=TRUE,
                 header = T, sep = ";", dec = ".", 
                 na.strings = c("", " ", "NA", "999", 999, "999.00", 999.00, 
                                "Missing", "missing"))

# variable selection
cvar <- c("Record_ID", "Readmission","HypoP", "surgery_type", "Sex", # "Indicatie", "Center", "Validatie"
          "BSKgezien",  "CHKD", "BSKinPA", "PostPA", "PostOPCa", "Validatie1")
nvar <- c("BaselinePTH", "PTH24u", "BaseCa", "Basealbu", "Ca24u", "Albu24u",
           "Age_Years")
work.data <- original.data[c(cvar, nvar)]
```

# Data cleaning
```{r, eval=TRUE}
# coding of variables
work.data$HypoP <- ifelse(work.data$HypoP %in% c("No ", "No"), "No", "Yes")
work.data$HypoP <- as.factor(work.data$HypoP)
work.data$PostOPCa <- ifelse(work.data$PostOPCa %in% c("No ", "No"), "No", "Yes")
work.data$PostOPCa <- as.factor(work.data$PostOPCa)
work.data$Age_Years <- as.numeric(work.data$Age_Years)

# replace , by .
work.data$BaselinePTH <- gsub(",", ".", work.data$BaselinePTH)
work.data$BaselinePTH <- as.numeric(work.data$BaselinePTH)
work.data$PTH24u <- gsub(",", ".", work.data$PTH24u)
work.data$PTH24u <- as.numeric(work.data$PTH24u)
work.data$BaseCa <- gsub(",", ".", work.data$BaseCa)
work.data$BaseCa <- as.numeric(work.data$BaseCa)
work.data$Basealbu <- gsub(",", ".", work.data$Basealbu)
work.data$Basealbu <- as.numeric(work.data$Basealbu)
work.data$Ca24u <- gsub(",", ".", work.data$Ca24u)
work.data$Ca24u <- as.numeric(work.data$Ca24u)
work.data$Albu24u <- gsub(",", ".", work.data$Albu24u)
work.data$Albu24u <- as.numeric(work.data$Albu24u)
```

# Imputation
```{r, eval=TRUE}
# single imputation is fine since less than 10% is missing
colMeans(is.na(work.data))*100
save(work.data, file=paste0(file.path, "/Data/data.to.be.imputed.Rdata"))

# set-up imputation
miceHypoP <- mice(work.data, maxit = 0)
miceHypoPmeth <- miceHypoP$meth
miceHypoPpred <- miceHypoP$pred
miceHypoP

## will not be used as predictor in model
miceHypoPpred[, "Record_ID"] <- 0

# single imputation
m <- 10
m.choose <- 1
mi <- mice::mice(work.data, m = m, seed = seed,
           method=miceHypoPmeth,
           predictorMatrix=miceHypoPpred,
           print=FALSE)
```

# Berekenen PTH daling, gecorrigeerd calcium en gecorrigeerd calcium daling
```{r, eval=TRUE}
# imputed data in long format
mi_long <- mice::complete(mi, action = "long", include = T)

# calculate delta PTH, delta Ca, and delta CorrCa
mi_long$dPTH <- ((mi_long$BaselinePTH - mi_long$PTH24u)/mi_long$BaselinePTH*100)
mi_long$CorrCaBaseline <- mi_long$BaseCa +((34-mi_long$Basealbu)*0.016)
mi_long$CorrCa24u <- mi_long$Ca24u +((34-mi_long$Albu24u)*0.016)
mi_long$dCorrCa24u <- ((mi_long$CorrCaBaseline - mi_long$CorrCa24u)/mi_long$CorrCaBaseline*100)
mi_long$dCa24u <- ((mi_long$BaseCa - mi_long$Ca24u)/mi_long$BaseCa*100)

# imputed data into mice format
imputed.data <- as.mids(mi_long)

# set data distributions
single.imputation <- mice::complete(imputed.data, m.choose)
dd<-rms::datadist(single.imputation)
options(datadist='dd')
options(digits=8)
```
# 1. Functional form - possible predictors
```{r, eval=TRUE, warning=FALSE}
# include either PTH24u or dPTH
# include either Ca24u or CorrCa24u or dCa24u or dCorrCa24u
functional.forms <- list(
  "PTH24u + Ca24u",
  "PTH24u + CorrCa24u",
  "PTH24u + dCa24u",
  "PTH24u + dCorrCa24u",
  "dPTH + Ca24u",
  "dPTH + CorrCa24u", # BEST and clinically most practical
  "dPTH + dCa24u",
  "dPTH + dCorrCa24u")
model.AICs <- c()
for (functional.form in functional.forms){
  form <- eval(parse(text=paste("HypoP ~", functional.form, "+ BSKgezien + Age_Years + Sex + surgery_type + CHKD")))
  model <- rms::lrm(form, data=single.imputation) # SINGLE IMPUTATION
  model.AICs <- c(model.AICs, stats::AIC(model))
}
possible.predictors.df <- as.data.frame(cbind(functional.forms, sprintf("%.1f", model.AICs)))
openxlsx::write.xlsx(possible.predictors.df,
            rowNames=FALSE,
            file=paste0(file.path, "/Results/functional.form.xlsx"))
print(cbind(functional.forms, model.AICs))
cat("Minimum AIC:", which(model.AICs==min(model.AICs)))

# Warning message means that we predict exact 0 and 1 due to smalls ample size (see https://www.statology.org/glm-fit-fitted-probabilities-numerically-0-or-1-occurred/)
```
# 2. Functional form - flexibility
```{r, eval=TRUE, warning=FALSE}
# plot most flexible model
form.flexible.model <- HypoP ~ rms::rcs(dPTH, 4) + 
  rms::rcs(CorrCa24u, 4) + 
  BSKgezien + 
  rms::rcs(Age_Years, 4) + 
  Sex + 
  surgery_type + 
  CHKD
flexible.model <- rms::lrm(form.flexible.model, data=single.imputation, 
                           maxit=1000, x=TRUE, y=TRUE)
grDevices::png(file=paste0(file.path, "Results/plot.Predict.png"))
plot(Predict(flexible.model))
grDevices::dev.off()

# check if dPTH, CorrCa or age need to be modelled non-linearly
functional.forms <- list(
  "dPTH + CorrCa24u + Age_Years", # RIGID MODEL
  "rms::rcs(dPTH, 3) + CorrCa24u + Age_Years", 
  "rms::rcs(dPTH, 4) + CorrCa24u + Age_Years", # BEST
  "rms::rcs(dPTH, 5) + CorrCa24u + Age_Years", 
  "dPTH + rms::rcs(CorrCa24u, 3) + Age_Years",
  "dPTH + rms::rcs(CorrCa24u, 4) + Age_Years",
  "dPTH + rms::rcs(CorrCa24u, 5) + Age_Years",
  "dPTH + CorrCa24u + rms::rcs(Age_Years, 3)", 
  "dPTH + CorrCa24u + rms::rcs(Age_Years, 4)", 
  "dPTH + CorrCa24u + rms::rcs(Age_Years, 5)", 
  "rms::rcs(dPTH, 4) + rms::rcs(CorrCa24u, 3) + Age_Years",
  "rms::rcs(dPTH, 4) + CorrCa24u + rms::rcs(Age_Years, 3)",
  "dPTH + rms::rcs(CorrCa24u, 3) + rms::rcs(Age_Years, 3)",
  "rms::rcs(dPTH, 4) + rms::rcs(CorrCa24u, 3) + rms::rcs(Age_Years, 3)") 
model.AICs <- c()
for (functional.form in functional.forms){
  form <- eval(parse(text=paste("HypoP ~", functional.form, "+ BSKgezien + Sex + surgery_type + CHKD")))
  model <- rms::lrm(form, data=single.imputation, # SINGLE IMPUTATION
                    maxit=1000, x=TRUE, y=TRUE)
  model.AICs <- c(model.AICs, stats::AIC(model))
}
flexibility.df <- as.data.frame(cbind(functional.forms, sprintf("%.1f", model.AICs)))
openxlsx::write.xlsx(flexibility.df,
            rowNames=FALSE,
            file=paste0(file.path, "/Results/flexibility.xlsx"))
print(cbind(functional.forms, model.AICs))
cat("Minimum AIC:", which(model.AICs==min(model.AICs)))
```
# Full model
```{r, eval=TRUE, warning=FALSE}
# linear
form.full.model <- HypoP ~ dPTH + 
  CorrCa24u + 
  BSKgezien + 
  Age_Years + 
  Sex +
  surgery_type +
  CHKD

# fit using multiple imputation
full.model <- Hmisc::fit.mult.impute(form.full.model, lrm, 
                                          xtrans=imputed.data,
                                          data=work.data,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE))
```

# Compare best flexible model with full model in single imputation
```{r, eval=TRUE}
full.model.single <- rms::lrm(form.full.model, data=single.imputation, # SINGLE IMPUTATION
                              x=TRUE, y=TRUE)
best.flexible.model <- rms::lrm(HypoP ~ rms::rcs(dPTH, 4) + CorrCa24u + BSKgezien + Age_Years + Sex + surgery_type + CHKD,
                                data=single.imputation, # SINGLE IMPUTATION
                                maxit=1000, x=TRUE, y=TRUE)
lrtest(best.flexible.model, full.model.single)
```

# Calculate optimism using bootstrap using full model and doing backward selection for each bootstrap
```{r, eval=TRUE, warning=FALSE}
# for single imputation
v <- rms::validate(full.model, method="boot", bw=TRUE, rule="p", sls=0.05, B=100,
                   pr=FALSE, type="individual")

# Show shrinkage factor
shrinkage.factor <- v["Slope","test"]
cat("Shrinkage factor     :", shrinkage.factor, "\n")
save(shrinkage.factor, work.data, imputed.data, single.imputation, 
     file=paste0(file.path, "/Data/test.Rdata"))

# Optimism corrected C-index
cindex.orig<-(v["Dxy","index.orig"]+1)/2   # original C-index
optimism.C<-v["Dxy","optimism"]/2          # optimism C-index, /2 because  C.orig - C.B the +1 cancels out
cindex.B<-(v["Dxy","index.corrected"]+1)/2 # index corrected C-index
cat(" Original C          :", cindex.orig, "\n",
    "Optimism            :", optimism.C, "\n",
    "Mean bootstrapped C :", cindex.B, "\n",
    "Optimism corrected C:", cindex.orig-optimism.C, "\n")
```

# Backward selection
```{r, eval=TRUE, warning=FALSE}
form.final.model <- HypoP ~ dPTH + 
  CorrCa24u +
  BSKgezien 

# multiple imputation model
final.model <- Hmisc::fit.mult.impute(form.final.model, lrm, 
                                          xtrans=imputed.data,
                                          data=work.data,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE))
```

# Simple model
```{r, eval=TRUE, warning=FALSE}
simple.model <- Hmisc::fit.mult.impute(HypoP ~ dPTH, lrm, 
                                          xtrans=imputed.data,
                                          data=work.data,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE))
```

# Save results of full model
```{r, eval=TRUE}
model.types <- c("full", "final", "simple")
for (model.type in model.types){
  # new intercept after shrinkage
  fitted.model <- eval(parse(text=paste0(model.type, ".model")))
  lp <- predict(fitted.model)
  lp.shrunk <- shrinkage.factor*lp
  fit.extra <- lrm.fit(y=single.imputation$HypoP,
                        offset=lp.shrunk) # TODO: for each imputation?
  intercept.extra <- coef(fit.extra)[1]
  intercept.shrunk <- shrinkage.factor*fitted.model$coefficients["Intercept"]+intercept.extra
  
  # check
  mean(single.imputation$HypoP=="Yes") # event probability
  mean(plogis(lp.shrunk+intercept.extra))  # same
  
  # save results
  if (model.type=="simple"){
    summary.model <- summary(fitted.model, 
                           dPTH=c(0, 1))[which(summary(fitted.model)[, "Type"]==1),]
    coef.model <- shrinkage.factor*summary.model["Effect"]
    CI.coef <- summary.model[c("Lower 0.95", "Upper 0.95")]
  } else{
    summary.model <- summary(fitted.model, 
                           dPTH=c(0, 1),
                           CorrCa24u=c(0.2, 0.1),
                           BSKgezien="Yes")[which(summary(fitted.model)[, "Type"]==1), ]
    coef.model <- shrinkage.factor*summary.model[,  "Effect"]
    CI.coef <- summary.model[, c("Lower 0.95", "Upper 0.95")]
  }
  
  # save results
  assign(paste0(model.type, ".intercept.shrunk"), intercept.shrunk)
  assign(paste0("coef.", model.type, ".model"), coef.model)
  
  an <- stats::anova(fitted.model)
  assign(paste0("an.", model.type), an)
  
  Int.CI.lower <- fitted.model$coefficients["Intercept"]+qnorm(0.025)*sqrt(diag(fitted.model$var)[1]/nrow(work.data))
  Int.CI.upper <- fitted.model$coefficients["Intercept"]+qnorm(0.975)*sqrt(diag(fitted.model$var)[1]/nrow(work.data))
  CI.df <- exp(rbind(c(shrinkage.factor*Int.CI.lower+intercept.extra,
                       shrinkage.factor*Int.CI.upper+intercept.extra),
                     shrinkage.factor*CI.coef))
  if (model.type!="simple"){
    row.names(CI.df)[1] <- c("Intercept") # give intercept a name
  } else{
    CI.df <- as.data.frame(CI.df)
    row.names(CI.df) <- c("Intercept", "dPTH")
  }
  assign(paste0("CI.", model.type, ".model"), CI.df)
  
  assign(paste0("OR.", model.type, ".model"), 
         exp(c(intercept.shrunk, coef.model)))
}

coef.final.model.webapp <- shrinkage.factor*summary(final.model, 
                                   dPTH=c(1, 2), 
                                   CorrCa24u=c(1, 2), 
                                   BSKgezien="No")[which(summary(final.model)[, "Type"]==1), 
                                                   c("Effect")]
save(shrinkage.factor, final.intercept.shrunk, coef.final.model.webapp, 
     file=paste0(file.path, "/Results/final.model.Rdata"))
```

# Assess performance
```{r, eval=TRUE, warning=FALSE}
# make predictions for each imputation
lp.simple <- c()
lp.final <- c()
lp.full <- c()
for (i in 1:m){
  lp.simple <- cbind(lp.simple, predict(simple.model, newdata=mice::complete(imputed.data, i))) # TODO: apply shrinkage factor?
  lp.final <- cbind(lp.final, predict(final.model, newdata=mice::complete(imputed.data, i)))
  lp.full <- cbind(lp.full, predict(full.model, newdata=mice::complete(imputed.data, i)))
}

# simple model
png(file=paste0(file.path, "Results/model.performance.simple.model.png"), 
    width=400, height=400, units="px")
out.simple <- PredictionTools::val.prob.mi(lp.mi=lp.simple, y=as.numeric(work.data$HypoP)-1, dist=TRUE)
grDevices::dev.off()
lrtest(simple.model, final.model)

# final model
png(file=paste0(file.path, "Results/model.performance.final.model.png"), 
    width=400, height=400, units="px")
out.final <- PredictionTools::val.prob.mi(lp.mi=lp.final, y=as.numeric(work.data$HypoP)-1, dist=TRUE)
grDevices::dev.off()
lrtest(final.model, full.model)

# full model
png(file=paste0(file.path, "Results/model.performance.full.model.png"), 
    width=400, height=400, units="px")
out.full <- PredictionTools::val.prob.mi(lp.mi=lp.full, y=as.numeric(work.data$HypoP)-1, dist=TRUE)
grDevices::dev.off()
```
# Save table
```{r, eval=TRUE, warning=FALSE}
names.coef <- c("Intercept", "dPTH", "CorrCa24u", "BSKgezien - No:Yes", 
                               "Age_Years", "Sex - Male:Female", 
                               "surgery_type - completion:total", "CHKD - Yes:No")
OR.table <- data.frame(names=names.coef,
                                coef.full=sprintf("%.2f", OR.full.model[names.coef]),
                                CI.full=c(paste0("[", sprintf("%.2f", CI.full.model[names.coef, 1]), "; ", 
                                               sprintf("%.2f", CI.full.model[names.coef, 2]), "]")),
                                Chi.full=c("", sprintf("%.1f", an.full[-nrow(an.full), "Chi-Square"])),
                                coef.final=c(sprintf("%.2f", OR.final.model), rep("", 4)),
                                CI.final=c(paste0("[", sprintf("%.2f", CI.final.model[, 1]), "; ", 
                                               sprintf("%.2f", CI.final.model[, 2]), "]"), rep("", 4)),
                                Chi.final=c("", sprintf("%.1f", an.final[-nrow(an.final), "Chi-Square"]),
                                            rep("", 4)),
                                coef.simple=c(sprintf("%.2f", OR.simple.model), rep("", 6)),
                                CI.simple=c(paste0("[", sprintf("%.2f", CI.simple.model[, 1]), "; ", 
                                               sprintf("%.2f", CI.simple.model[, 2]), "]"),
                                            rep("", 6)),
                                Chi.simple=c("", sprintf("%.1f", an.simple[-nrow(an.simple), "Chi-Square"]),
                                             rep("", 6)))
Cindex.table <- c("C-index",
                  sprintf("%.2f", out.full$cindex-optimism.C), 
                  paste0("[", sprintf("%.2f", out.full$cindex.lower-optimism.C), "; ", sprintf("%.2f", out.full$cindex.upper-optimism.C), "]"),
                  "",
                  sprintf("%.2f", out.final$cindex-optimism.C), 
                  paste0("[", sprintf("%.2f", out.final$cindex.lower-optimism.C), "; ", sprintf("%.2f", out.final$cindex.upper-optimism.C), "]"),
                  "",
                  sprintf("%.2f", out.simple$cindex-optimism.C), 
                  paste0("[", sprintf("%.2f", out.simple$cindex.lower-optimism.C), "; ", sprintf("%.2f", out.simple$cindex.upper-optimism.C), "]"),
                  "")
combined.table <- rbind(OR.table, Cindex.table)
final.table <- cbind(combined.table[, 1:4], rep("", nrow(combined.table)),
                     combined.table[, 5:7], rep("", nrow(combined.table)),
                     combined.table[, 8:10])
openxlsx::write.xlsx(final.table,
            rowNames=FALSE,
            file=paste0(file.path, "/Results/model.xlsx"))

```
# Assess leave-one-center-out cross-validation
```{r, eval=TRUE, warning=FALSE}
# 4 splits
table(work.data$Validatie1)
Centers <- levels(work.data$Validatie1)
for (type in c("full", "final")){
  for (j in 1:4){
    # split outcomes for center j and not j
    y.j <- as.numeric(work.data$HypoP[work.data$Validatie1==Centers[j]])-1
    y.notj <- as.numeric(work.data$HypoP[work.data$Validatie1!=Centers[j]])-1
    
    # model form for center j and not j
    form.model <- eval(parse(text=paste0("form.", type, ".model")))
    form.notj <- update(form.model, y.notj  ~ . )
    
    # fit on 3 centers
    model.notj <- rms::lrm(form.notj, data=single.imputation[work.data$Validatie1!=Centers[j],], x=TRUE, y=TRUE)
    
    # TODO: calculate shrinkage?
    # v<-rms::validate(model.notj,method="boot",bw=TRUE,rule="p",sls=0.05,B=100,pr=FALSE,type="individual")
    # shrinkage.factor.notj <- v["Slope","test"]
    
    # cat("Now calculating for center", Centers[j], "\n")
    # print(model.notj$coefficients)
    # print(shrinkage.factor.notj)
    
    # make prediction for 1 center
    lp.j <- predict(model.notj, newdata=single.imputation[work.data$Validatie1==Centers[j],])

    # make plot
    png(file=paste0(file.path, "Results/Figures/model.performance.", type, ".", Centers[j], ".png"))
    out.j <- PredictionTools::val.prob.mi(lp.mi=lp.j, y=y.j, main=Centers[j], g=3, dist=TRUE)
    grDevices::dev.off()
  }
}

# combine 4 calibration plots into one
for (type in c("full", "final")){
  png(file=paste0(file.path, "Results/leave.one.out.cross.validation.", type, ".png"), width=16, height=16, units="cm", res=300)
  par(mar=rep(0, 4))
  layout(matrix(1:4, ncol=2, byrow=TRUE))
  Centers.val <- rep(1:4, 3)
  for (i in 1:4){
    plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
    img <- png::readPNG(paste0(file.path, "Results/Figures/model.performance.", type, ".", Centers[Centers.val[i]], ".png"))
    rasterImage(img, 0, 0, 1, 1)
  }
  grDevices::dev.off()
}
```
# dPTH and readmission
```{r, eval=TRUE}
dPTH <- c()
lp <- c()
for (i in 1:m){
  selected.imputation <- mice::complete(imputed.data, i)
  
  dPTH <- cbind(dPTH, selected.imputation$dPTH)
  
  lp <- cbind(lp,
              final.intercept.shrunk+
                  coef.final.model.webapp["dPTH"]*selected.imputation$dPTH+
                  coef.final.model.webapp["CorrCa24u"]*selected.imputation$CorrCa24u+
                  coef.final.model.webapp["BSKgezien - Yes:No"]*ifelse(selected.imputation$BSKgezien == "Yes", 1, 0))
}
# if the majority of imputations gave high dPTH, consider this patient high dPTH
table(single.imputation$Readmission, rowMeans(dPTH)>0.7)

# average predicted probability of all imputations
pred <- rowMeans(exp(lp)/(1+exp(lp)))

# predictions stratified
str.pred <- ifelse(pred<0.1, 1, ifelse(pred>0.3, 3, 2))
table(str.pred)
# absolute numbers
table(single.imputation$Readmission, str.pred)
# percentages
round(table(single.imputation$Readmission, str.pred)/rbind(table(str.pred), table(str.pred))*100, 1)
```
