---
title: "Prediction model for permanent hypoparathyroidism"
author: "Carolien C.H.M. Maas, Erasmus MC, Rotterdam, The Netherlands"
date: "July 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages, functions, and data
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
seed <- 1
set.seed(seed)

# load libraries
library(mice)
library(rms)
# remotes::install_github("CHMMaas/PredictionTools")
library(PredictionTools)
source("Z:/Project Predict Hypoparathyroidism/Development/Code/propplot_functie.R")
```

# Load data
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
file.path <- "Z:/Project Predict Hypoparathyroidism/Development/"
original.data <- utils::read.csv(paste0(file.path, "Data/Data.csv"), 
                 stringsAsFactors=TRUE,
                 header = T, sep = ";", dec = ".", 
                 na.strings = c("", " ", "NA", "999", 999, "999.00", 999.00, 
                                "Missing", "missing"))
new.imputation <- FALSE

# variable selection
cvar <- c("Record_ID", "Readmission","HypoP", "surgery_type", "Sex", # "Indicatie", "Center", "Validatie"
          "BSKgezien",  "CHKD", "BSKinPA", "PostPA", "PostOPCa", "Validatie1")
nvar <- c("BaselinePTH", "PTH24u", "BaseCa", "Basealbu", "Ca24u", "Albu24u",
           "Age_Years")
work.data <- original.data[c(cvar, nvar)]
```

# Data cleaning
```{r, eval=TRUE}
# coding of variables
work.data$HypoP <- ifelse(work.data$HypoP %in% c("No ", "No"), "No", "Yes")
work.data$HypoP <- as.factor(work.data$HypoP)
work.data$PostOPCa <- ifelse(work.data$PostOPCa %in% c("No ", "No"), "No", "Yes")
work.data$PostOPCa <- as.factor(work.data$PostOPCa)
work.data$Age_Years <- as.numeric(work.data$Age_Years)

# replace , by .
work.data$BaselinePTH <- gsub(",", ".", work.data$BaselinePTH)
work.data$BaselinePTH <- as.numeric(work.data$BaselinePTH)
work.data$PTH24u <- gsub(",", ".", work.data$PTH24u)
work.data$PTH24u <- as.numeric(work.data$PTH24u)
work.data$BaseCa <- gsub(",", ".", work.data$BaseCa)
work.data$BaseCa <- as.numeric(work.data$BaseCa)
work.data$Basealbu <- gsub(",", ".", work.data$Basealbu)
work.data$Basealbu <- as.numeric(work.data$Basealbu)
work.data$Ca24u <- gsub(",", ".", work.data$Ca24u)
work.data$Ca24u <- as.numeric(work.data$Ca24u)
work.data$Albu24u <- gsub(",", ".", work.data$Albu24u)
work.data$Albu24u <- as.numeric(work.data$Albu24u)
```

# Imputation
```{r, eval=TRUE}
# single imputation is fine since less than 10% is missing
colMeans(is.na(work.data))*100
save(work.data, file=paste0(file.path, "/Data/data.to.be.imputed.Rdata"))

# set-up imputation
miceHypoP <- mice(work.data, maxit = 0)
miceHypoPmeth <- miceHypoP$meth
miceHypoPpred <- miceHypoP$pred
miceHypoP

## will not be used as predictor in model
miceHypoPpred[, "Record_ID"] <- 0

# single imputation
m <- 10
m.choose <- 1
if (new.imputation){
  mi <- mice::mice(work.data, m = m, seed = seed,
           method=miceHypoPmeth,
           predictorMatrix=miceHypoPpred,
           print=FALSE)
  save(mi, file=paste0(file.path, "/Data/imputed.data.Rdata"))
} else{
  load(paste0(file.path, "/Data/imputed.data.Rdata"))
}

# plots
propplot(mi)
mice::densityplot(mi)
```

# Berekenen PTH daling, gecorrigeerd calcium en gecorrigeerd calcium daling
```{r, eval=TRUE}
# imputed data in long format
mi_long <- mice::complete(mi, action = "long", include = T)

# calculate delta PTH, delta Ca, and delta CorrCa
mi_long$dPTH <- ((mi_long$BaselinePTH - mi_long$PTH24u)/mi_long$BaselinePTH*100)
mi_long$CorrCaBaseline <- mi_long$BaseCa +((34-mi_long$Basealbu)*0.016)
mi_long$CorrCa24u <- mi_long$Ca24u +((34-mi_long$Albu24u)*0.016)
mi_long$dCorrCa <- ((mi_long$CorrCaBaseline - mi_long$CorrCa24u)/mi_long$CorrCaBaseline*100)
mi_long$dCa <- ((mi_long$BaseCa - mi_long$Ca24u)/mi_long$BaseCa*100)

# imputed data into mice format
imputed.data <- as.mids(mi_long)

# set data distributions
single.imputation <- mice::complete(imputed.data, m.choose)
dd<-rms::datadist(single.imputation)
options(datadist='dd')
options(digits=8)
```
# Full model
```{r, eval=TRUE, warning=FALSE}
# linear
form.full.model <- HypoP ~ dPTH + 
  BSKgezien + 
  CorrCa24u + 
  dCorrCa +
  Age_Years + 
  Sex +
  surgery_type +
  CHKD

# fit using multiple imputation
full.model <- Hmisc::fit.mult.impute(form.full.model, lrm, 
                                          xtrans=imputed.data,
                                          data=work.data,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE))
```

# Calculate optimism using bootstrap using full model and doing backward selection for each bootstrap
```{r, eval=TRUE, warning=FALSE}
# for single imputation
v <- rms::validate(full.model, method="boot", bw=TRUE, rule="p", sls=0.05, B=1000,
                   pr=FALSE, type="individual")

# Show shrinkage factor
shrinkage.factor.hypop <- v["Slope","test"]
cat("Shrinkage factor     :", shrinkage.factor.hypop, "\n")

# Optimism corrected C-index
cindex.orig<-(v["Dxy","index.orig"]+1)/2   # original C-index
optimism.C<-v["Dxy","optimism"]/2          # optimism C-index, /2 because  C.orig - C.B the +1 cancels out
cindex.B<-(v["Dxy","index.corrected"]+1)/2 # index corrected C-index
cat(" Original C          :", cindex.orig, "\n",
    "Optimism            :", optimism.C, "\n",
    "Mean bootstrapped C :", cindex.B, "\n",
    "Optimism corrected C:", cindex.orig-optimism.C, "\n")
```

# Backward selection
```{r, eval=TRUE, warning=FALSE}
form.final.model <- HypoP ~ dPTH + 
  BSKgezien +
  CorrCa24u

# multiple imputation model
final.model <- Hmisc::fit.mult.impute(form.final.model, lrm, 
                                          xtrans=imputed.data,
                                          data=work.data,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE))
```
# Save results of full and final model
```{r, eval=TRUE}
model.types <- c("full", "final")
for (model.type in model.types){
  # new intercept after shrinkage
  fitted.model <- eval(parse(text=paste0(model.type, ".model")))
  intercepts.extra <- c()
  lp.shrunk <- c()
  for (i in 1:m){
    lp.i <- predict(fitted.model, newdata=mice::complete(imputed.data, i))
    lp.shrunk.i <- shrinkage.factor.hypop*lp.i
    lp.shrunk <- cbind(lp.shrunk, lp.shrunk.i)
    fit.extra.i <- lrm.fit(y=work.data$HypoP,
                           offset=lp.shrunk.i)
    intercepts.extra <- c(intercepts.extra, coef(fit.extra.i)[1])
  }
  intercept.extra <- mean(intercepts.extra)
  intercept.shrunk <- shrinkage.factor.hypop*fitted.model$coefficients["Intercept"]+intercept.extra
  
  # check
  mean(work.data$HypoP=="Yes") # event probability
  mean(plogis(rowMeans(lp.shrunk)+intercept.extra))  # same
  
  # save results
  if (model.type=="full"){
    summary.model <- summary(fitted.model, 
                         dPTH=c(0, 1),
                         CorrCa24u=c(0.2, 0.1),
                         dCorrCa=c(2, 1),
                         BSKgezien="Yes")[which(summary(fitted.model)[, "Type"]==1), ]
  } else{
    summary.model <- summary(fitted.model, 
                         dPTH=c(0, 1),
                         CorrCa24u=c(0.2, 0.1),
                         BSKgezien="Yes")[which(summary(fitted.model)[, "Type"]==1), ]
  }
  
  coef.model <- c(fitted.model$coefficients["Intercept"], summary.model[,  "Effect"])
  CI.coef <- summary.model[, c("Lower 0.95", "Upper 0.95")]
  shrunk.coef.model <- shrinkage.factor.hypop*summary.model[,  "Effect"]

  Int.CI.lower <- fitted.model$coefficients["Intercept"]+qnorm(0.025)*sqrt(diag(fitted.model$var)[1]/nrow(work.data))
  Int.CI.upper <- fitted.model$coefficients["Intercept"]+qnorm(0.975)*sqrt(diag(fitted.model$var)[1]/nrow(work.data))
  CI.df <- exp(rbind(c(Int.CI.lower, Int.CI.upper), CI.coef))
  
  # save results
  assign(paste0("coef.", model.type, ".model"), coef.model)
  assign(paste0("OR.", model.type, ".model"), exp(coef.model))
  assign(paste0("lp.shrunk.", model.type), lp.shrunk)
  assign(paste0("shrunk.intercept.", model.type), intercept.shrunk)
  assign(paste0("shrunk.coef.", model.type, ".model"), shrunk.coef.model)
  assign(paste0("shrunk.OR.", model.type, ".model"), exp(c(intercept.shrunk, shrunk.coef.model)))
  
  an <- stats::anova(fitted.model)
  assign(paste0("an.", model.type), an)
  
  row.names(CI.df)[1] <- c("Intercept") # give intercept a name
  assign(paste0("CI.", model.type, ".model"), CI.df)
}
```

# Assess performance (without shrinkage)
```{r, eval=TRUE, warning=FALSE}
for (model.type in c("full", "final")){
  # make predictions for each imputation
  lp.model <- c()
  fitted.model <- eval(parse(text=paste0(model.type, ".model")))
  for (i in 1:m){
    lp.model <- cbind(lp.model, predict(fitted.model, 
                                        newdata=mice::complete(imputed.data, i)))
  }
  assign(paste0("lp.", model.type), lp.model)
  
  # make plot
  png(file=paste0(file.path, "Results/Figures/model.performance.", model.type, ".model.png"), 
      width=400, height=400, units="px")
  out.model <- PredictionTools::val.prob.mi(lp.mi=lp.model,
                               y=as.numeric(work.data$HypoP)-1, 
                               g=3, dist=TRUE, lim=c(0, 1),
                               show.metrics=c(TRUE, TRUE, rep(FALSE, 3), TRUE, rep(FALSE, 3)))
  grDevices::dev.off()
  assign(paste0("out.", model.type), out.model)
}
lrtest(final.model, full.model)
```
# Assess leave-one-cohort-out cross-validation (without shrinkage)
```{r, eval=TRUE, warning=FALSE}
# 4 splits
table(work.data$Validatie1)
Centers <- levels(work.data$Validatie1)
Centers.title <- paste(c("A.", "B.", "C.", "D."), "Cohort", 1:4)
for (type in c("full", "final")){
  for (j in 1:4){
    # split outcomes for center j and not j
    y.hypop.j <- as.numeric(work.data$HypoP[work.data$Validatie1==Centers[j]])-1         # test
    y.read.j <- as.numeric(work.data$Readmission[work.data$Validatie1==Centers[j]])-1    # test
    y.hypop.notj <- as.numeric(work.data$HypoP[work.data$Validatie1!=Centers[j]])-1      # train
    y.read.notj <- as.numeric(work.data$Readmission[work.data$Validatie1!=Centers[j]])-1 # train
    
    # model form for center j and not j
    form.model <- eval(parse(text=paste0("form.", type, ".model")))
    form.notj <- update(form.model, y.hypop.notj  ~ . )
    
    # fit on 3 centers
    model.notj <- Hmisc::fit.mult.impute(form.notj, lrm, 
                                          xtrans=imputed.data,
                                          data=work.data,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE),
                                          sub=work.data$Validatie1!=Centers[j])
    
    # make prediction for 1 center
    lp.hypop.i <- c()
    lp.cal.read.i <- c()
    for (i in 1:m){
      # linear predictor for hypoparathyroidism
      data.i <- mice::complete(imputed.data, i)
      lp.hypop.i <- cbind(lp.hypop.i, predict(model.notj, newdata=data.i[work.data$Validatie1==Centers[j],]))
      
      # calibrate model for readmission
      lp.notj <- predict(model.notj, newdata=data.i[work.data$Validatie1!=Centers[j],])
      f.notj <- rms::lrm(y.read.notj ~ lp.notj,
                         data=data.i[work.data$Validatie1!=Centers[j],],
                         y=TRUE, x=TRUE)
      cal.int.i <- stats::coef(f.notj)[1]
      cal.fact.i <- stats::coef(f.notj)[2]
      
      # linear predictor for calibrated model
      coef.calibrated.model.i <- c(cal.fact.i*coef(final.model)[1]+cal.int.i,
                                   cal.fact.i*coef(final.model)[-1])
      predict.df <- as.matrix(cbind(rep(1, nrow(data.i[work.data$Validatie1==Centers[j],])), # Intercept 
                                data.i[work.data$Validatie1==Centers[j], "dPTH"],            # dPTH
                                as.numeric(data.i[work.data$Validatie1==Centers[j], "BSKgezien"])-1, # BSKgezien
                                data.i[work.data$Validatie1==Centers[j], "CorrCa24u"]))   # CorrCa24u
      lp.cal.read.i <- cbind(lp.cal.read.i, as.numeric(t(coef.calibrated.model.i) %*% t(predict.df)))
    }
    
    # make plot hypoparathyroidism
    png(file=paste0(file.path, "Results/Figures/model.performance.hypop.", type, ".", Centers[j], ".png"))
    PredictionTools::val.prob.mi(lp.mi=lp.hypop.i, y=y.hypop.j, main=Centers.title[j], 
                                  g=3, dist=TRUE, smoothed.curve=FALSE, lim=c(0, 1),
                                  show.metrics=c(TRUE, TRUE, rep(FALSE, 3), TRUE, rep(FALSE, 3)))
    grDevices::dev.off()
    
    # make plot readmission
    png(file=paste0(file.path, "Results/Figures/model.performance.readmission.", type, ".", Centers[j], ".png"))
    PredictionTools::val.prob.mi(lp.mi=lp.cal.read.i, y=y.read.j, main=Centers.title[j], 
                                  g=3, dist=TRUE, smoothed.curve=FALSE, lim=c(0, 1),
                                  show.metrics=c(TRUE, TRUE, rep(FALSE, 3), TRUE, rep(FALSE, 3)))
    grDevices::dev.off()
  }
}

# combine 4 calibration plots into one
for (type in c("full", "final")){
  png(file=paste0(file.path, "Results/", ifelse(type=="full", "Figures/", ""),
                  "leave.one.out.cross.validation.hypop.", type, ".png"), width=16, height=16, units="cm", res=300)
  par(mar=rep(0, 4))
  layout(matrix(1:4, ncol=2, byrow=TRUE))
  Centers.val <- rep(1:4, 3)
  for (i in 1:4){
    plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
    img <- png::readPNG(paste0(file.path, "Results/Figures/model.performance.hypop.", type, ".", Centers[Centers.val[i]], ".png"))
    rasterImage(img, 0, 0, 1, 1)
  }
  grDevices::dev.off()
  
  png(file=paste0(file.path, "Results/", ifelse(type=="full", "Figures/", ""),
                  "leave.one.out.cross.validation.readmission.", type, ".png"), width=16, height=16, units="cm", res=300)
  par(mar=rep(0, 4))
  layout(matrix(1:4, ncol=2, byrow=TRUE))
  Centers.val <- rep(1:4, 3)
  for (i in 1:4){
    plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
    img <- png::readPNG(paste0(file.path, "Results/Figures/model.performance.readmission.", type, ".", Centers[Centers.val[i]], ".png"))
    rasterImage(img, 0, 0, 1, 1)
  }
  grDevices::dev.off()
}
```
# Calibrated readmission model
```{r, eval=TRUE}
cal.int <- c()
cal.fact <- c()
cal.fact.SE <- c()
gammas.extra <- c()
for (i in 1:m){
  # calculate calibration factor
  lp.i <- predict(final.model,
                  newdata=mice::complete(imputed.data, i), 
                  type="lp")
  f.mi <- rms::lrm(Readmission~lp.i,
                   data=mice::complete(imputed.data, i),
                   y=TRUE, x=TRUE)
  cal.int <- c(cal.int, stats::coef(f.mi)[1])
  cal.fact <- c(cal.fact, stats::coef(f.mi)[2])
  cal.fact.SE <- c(cal.fact.SE, sqrt(diag(vcov(f.mi)))[2])
  
  # calculate correction for calibrated shrunk model
  lp.shrunk.i <- shrinkage.factor.hypop*lp.i
  fit.extra.i <- lrm.fit(y=work.data$Readmission,
                        offset=lp.shrunk.i)
  gammas.extra <- c(gammas.extra, coef(fit.extra.i)[1])
}
# factor added to intercept for calibrated shrunk coefficients
gamma.extra <- mean(gammas.extra)

# calibration factor is not significant from zero
cal.fact.combined <- PredictionTools::Rubin.combine(cal.fact, cal.fact.SE)
gamma <- cal.fact.combined$est
cat("Calibration factor:", paste0(sprintf("%.2f", gamma), " [",
            sprintf("%.2f", gamma+qnorm(.025)*cal.fact.combined$se), "; ",
            sprintf("%.2f", gamma+qnorm(.975)*cal.fact.combined$se), "]"))
```
# Calibration plots of calibrated readmission model (without shrinkage)
```{r, eval=TRUE}
lp.calibrated <- c()
for (i in 1:m){
  imputed.data.i <- mice::complete(imputed.data, i)
  predict.df <- as.matrix(cbind(rep(1, nrow(imputed.data.i)),
                                imputed.data.i[, "dPTH"],                    # dPTH
                                as.numeric(imputed.data.i[, "BSKgezien"])-1, # BSKgezien
                                imputed.data.i[, "CorrCa24u"]))              # CorrCa24u
  
  coef.calibrated.model.i <- c(cal.fact[i]*coef(final.model)[1]+cal.int[i],
                               cal.fact[i]*coef(final.model)[-1])
  lp.calibrated <- cbind(lp.calibrated, as.numeric(t(coef.calibrated.model.i) %*% t(predict.df)))
}

png(file=paste0(file.path, "Results/Figures/model.performance.calibrated.model.png"), 
    width=400, height=400, units="px")
out.calibrated <- PredictionTools::val.prob.mi(lp.mi=lp.calibrated, g=3, dist=TRUE, 
                           y=as.numeric(work.data$Readmission)-1, lim=c(0, 1),
                           show.metrics=c(TRUE, TRUE, rep(FALSE, 3), TRUE, rep(FALSE, 3)))
dev.off()
```
# Save Table
```{r, eval=TRUE, warning=FALSE}
names.coef.full <- c("Intercept", "dPTH", "BSKgezien - No:Yes", 
                "CorrCa24u", "dCorrCa",
                "Age_Years", "Sex - Male:Female",
                "surgery_type - completion:total", "CHKD - Yes:No")
names.coef.final <- c("Intercept", "dPTH", "BSKgezien - No:Yes", "CorrCa24u")
calibrated.readmission.coef <- c(gamma*shrunk.intercept.final+gamma.extra, 
                                 gamma*shrunk.coef.final.model)
calibrated.readmission.OR <- exp(calibrated.readmission.coef)
OR.table <- data.frame(names=names.coef.full,
              OR.full=sprintf("%.2f", OR.full.model[names.coef.full]),
              CI.full=c(paste0("[", sprintf("%.2f", CI.full.model[names.coef.full, 1]), "; ", 
                             sprintf("%.2f", CI.full.model[names.coef.full, 2]), "]")),
              shrunk.OR.full=sprintf("%.2f", shrunk.OR.full.model[names.coef.full]),
              Chi.full=c("", sprintf("%.1f", an.full[-nrow(an.full), "Chi-Square"])),
              OR.final=c(sprintf("%.2f", OR.final.model[names.coef.final]), rep("", 5)),
              CI.final=c(paste0("[", sprintf("%.2f", CI.final.model[names.coef.final, 1]), "; ", 
                             sprintf("%.2f", CI.final.model[names.coef.final, 2]), "]"), rep("", 5)),
              shrunk.OR.final=c(sprintf("%.2f", shrunk.OR.final.model[names.coef.final]), rep("", 5)),
              Chi.final=c("", sprintf("%.1f", an.final[-nrow(an.final), "Chi-Square"]),
                          rep("", 5)),
              calibrated.readmission=c(sprintf("%.2f", calibrated.readmission.OR[names.coef.final]), rep("", 5)))
Cindex.table <- c("C-index optimism-corrected",
                  sprintf("%.2f", out.full$cindex-optimism.C), 
                  paste0("[", sprintf("%.2f", out.full$cindex.lower-optimism.C), "; ", 
                         sprintf("%.2f", out.full$cindex.upper-optimism.C), "]"),
                  "",
                  "",
                  sprintf("%.2f", out.final$cindex-optimism.C), 
                  paste0("[", sprintf("%.2f", out.final$cindex.lower-optimism.C), "; ", 
                         sprintf("%.2f", out.final$cindex.upper-optimism.C), "]"),
                  "",
                  "",
                  paste0(sprintf("%.2f", out.calibrated$cindex-optimism.C), 
                         " [", sprintf("%.2f", out.calibrated$cindex.lower-optimism.C), "; ", 
                         sprintf("%.2f", out.calibrated$cindex.upper-optimism.C), "]"))
final.table <- rbind(OR.table, Cindex.table)
openxlsx::write.xlsx(final.table,
            rowNames=FALSE,
            file=paste0(file.path, "/Results/hypopara.model.xlsx"))

```
# Save for webapp
```{r, eval=TRUE}
coef.hypop.model.webapp <- c(shrunk.intercept.final,
                             shrinkage.factor.hypop*summary(final.model, 
                                   dPTH=c(1, 2), 
                                   CorrCa24u=c(1, 2), 
                                   BSKgezien="No")[which(summary(final.model)[, "Type"]==1), 
                                                   c("Effect")])
coef.readmission.model.webapp <- c(gamma*shrunk.intercept.final+mean(gamma.extra),
                                   gamma*coef.hypop.model.webapp[-1])

save(coef.hypop.model.webapp,
     coef.readmission.model.webapp,
     file=paste0(file.path, "/Results/final.model.Rdata"))
```
# dPTH and readmission
```{r, eval=TRUE}
dPTH <- c()
lp.hypop <- c()
lp.read <- c()
for (i in 1:m){
  selected.imputation <- mice::complete(imputed.data, i)

  dPTH <- cbind(dPTH, selected.imputation$dPTH)

  lp.hypop <- cbind(lp.hypop,
                  coef.hypop.model.webapp["Intercept"]+
                  coef.hypop.model.webapp["dPTH"]*selected.imputation$dPTH+
                  coef.hypop.model.webapp["CorrCa24u"]*selected.imputation$CorrCa24u+
                  coef.hypop.model.webapp["BSKgezien - Yes:No"]*ifelse(selected.imputation$BSKgezien == "Yes", 1, 0))
  
  lp.read <- cbind(lp.read,
                  coef.readmission.model.webapp["Intercept"]+
                  coef.readmission.model.webapp["dPTH"]*selected.imputation$dPTH+
                  coef.readmission.model.webapp["CorrCa24u"]*selected.imputation$CorrCa24u+
                  coef.readmission.model.webapp["BSKgezien - Yes:No"]*ifelse(selected.imputation$BSKgezien == "Yes", 1, 0))
}

# Supplemental Table 4
table(rowMeans(dPTH)>=70)
paste0(as.numeric(table(work.data$HypoP, rowMeans(dPTH)>=70)[2, ]), " (",
      round(table(work.data$HypoP, rowMeans(dPTH)>=70)[2, ]/table(rowMeans(dPTH)>=70)*100, 1), "%)")
paste0(as.numeric(table(work.data$Readmission, rowMeans(dPTH)>=70)[2, ]), " (",
      round(table(work.data$Readmission, rowMeans(dPTH)>=70)[2, ]/table(rowMeans(dPTH)>=70)*100, 1), "%)")

# Table 3 - predictions of hypoparathyroidism and readmissions
pred.hypop <- rowMeans(exp(lp.hypop)/(1+exp(lp.hypop)))
strat.pred.hypop <- ifelse(pred.hypop<0.1, 1, ifelse(pred.hypop>0.3, 3, 2))
table(strat.pred.hypop)
paste0(as.numeric(table(work.data$HypoP, strat.pred.hypop)[2, ]), " (",
       round(table(work.data$HypoP, strat.pred.hypop)/rbind(table(strat.pred.hypop), table(strat.pred.hypop))*100, 1)[2, ], "%)")
paste0(as.numeric(table(work.data$Readmission, strat.pred.hypop)[2, ]), " (",
       round(table(work.data$Readmission, strat.pred.hypop)/rbind(table(strat.pred.hypop), table(strat.pred.hypop))*100, 1)[2, ], "%)")
```
```{r, eval=TRUE}
PredictionTools::val.prob.mi(lp.mi=lp.hypop, y=as.numeric(work.data$HypoP)-1, dist=TRUE)
PredictionTools::val.prob.mi(lp.mi=lp.read, y=as.numeric(work.data$Readmission)-1, dist=TRUE)
```
